---
title: "STOR 565 Project Outline"
output: pdf_document
date: "2023-12-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Introduction

## Description of Project
### We are fitting an xG model
### Describe the usefulness of xG
### Give an overview of the process required to create an xG model


# Exploratory Data Analysis

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Import the tools necessary to use StatsBomb data
library(tidyverse)
library(devtools)
library(ggplot2)
library(StatsBombR)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Show the available selection of StatsBomb data (note that not all rows denote full seasons of data).
FreeCompetitions()
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Pull match data for the 2015/16 La Liga and Premier League seasons.

# Pull La Liga Match Data
Comp <- FreeCompetitions()[c(41),]
Matches <- FreeMatches(Comp)
StatsBombData <- free_allevents(MatchesDF = Matches, Parallel = T)
StatsBombData = allclean(StatsBombData)
laligadata = StatsBombData

# Pull Premier League Match Data
Comp <- FreeCompetitions()[c(59),]
Matches <- FreeMatches(Comp)
StatsBombData <- free_allevents(MatchesDF = Matches, Parallel = T)
StatsBombData = allclean(StatsBombData)
premdata = StatsBombData
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Create a dataframe that contains the shot event data for the seasons, as xG (expected goals) is a measure of shot/opportunity quality.

laligashotsdf = laligadata[laligadata$type.id==16,]
premshotsdf = premdata[premdata$type.id==16,]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
boollist = colnames(premshotsdf) %in% colnames(laligashotsdf)
boollist2 = colnames(laligashotsdf) %in% colnames(premshotsdf)
drop = colnames(premshotsdf)[!boollist]
drop2 = colnames(laligashotsdf)[!boollist2]

premtest = premshotsdf[,!(names(premshotsdf) %in% drop)]
laligatest = laligashotsdf[,!(names(laligashotsdf) %in% drop2)]

shotsdf = rbind(premtest, laligatest)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Drop irrelevant columns from our dataset (i.e. columns pertaining to non-shot related events like substitutions or passing).


cleanshotsdf = shotsdf[c("location", "under_pressure", "type.name", "possession_team.name", "play_pattern.name", "team.id", "team.name", "player.id", "player.name", "position.name", "shot.statsbomb_xg", "shot.freeze_frame", "shot.first_time", "shot.one_on_one", "shot.aerial_won", "shot.technique.name", "shot.body_part.name", "shot.outcome.name", "shot.type.name", "match_id", "shot.open_goal", "shot.follows_dribble", "location.x", "location.y", "location.x.GK", "location.y.GK", "DistToGoal", "DistToKeeper", "AngleToGoal", "AngleToKeeper", "AngleDeviation", "DistSGK", "density", "density.incone", "distance.ToD1", "distance.ToD2", "AttackersBehindBall", "DefendersBehindBall", "DefendersInCone", "InCone.GK", "DefArea", "distance.ToD1.360", "distance.ToD2.360")]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Drop rows that contain shots taken from penalty and free kicks, as these would need to be modeled as part of a separate exercise.
cleanshotsdf <- cleanshotsdf[cleanshotsdf$shot.type.name=="Open Play",]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Convert data to regression-friendly format

cleanshotsdf$under_pressure <- ifelse(is.na(cleanshotsdf$under_pressure), 0, 1)

cleanshotsdf$shot.first_time <- ifelse(is.na(cleanshotsdf$shot.first_time), 0, 1)

cleanshotsdf$shot.one_on_one <- ifelse(is.na(cleanshotsdf$shot.one_on_one), 0, 1)

cleanshotsdf$shot.aerial_won <- ifelse(is.na(cleanshotsdf$shot.aerial_won), 0, 1)

cleanshotsdf$shot.technique.normal <- ifelse((cleanshotsdf$shot.technique.name == "Normal"), 1, 0)
cleanshotsdf$shot.technique.halfvolley <- ifelse((cleanshotsdf$shot.technique.name == "Half Volley"), 1, 0)
cleanshotsdf$shot.technique.backheel <- ifelse((cleanshotsdf$shot.technique.name == "Backheel"), 1, 0)
cleanshotsdf$shot.technique.volley <- ifelse((cleanshotsdf$shot.technique.name == "Volley"), 1, 0)
cleanshotsdf$shot.technique.overheadkick <- ifelse((cleanshotsdf$shot.technique.name == "Overhead Kick"), 1, 0)
cleanshotsdf$shot.technique.divingheader <- ifelse((cleanshotsdf$shot.technique.name == "Diving Header"), 1, 0)
cleanshotsdf$shot.technique.lob <- ifelse((cleanshotsdf$shot.technique.name == "Lob"), 1, 0)


cleanshotsdf$shot.body_part.leftfoot <- ifelse((cleanshotsdf$shot.body_part.name == "Left Foot"), 1, 0)
cleanshotsdf$shot.body_part.rightfoot <- ifelse((cleanshotsdf$shot.body_part.name == "Right Foot"), 1, 0)
cleanshotsdf$shot.body_part.head <- ifelse((cleanshotsdf$shot.body_part.name == "Head"), 1, 0)
cleanshotsdf$shot.body_part.other <- ifelse((cleanshotsdf$shot.body_part.name == "Other"), 1, 0)


cleanshotsdf$shot.open_goal <- ifelse(is.na(cleanshotsdf$shot.open_goal), 0, 1)

cleanshotsdf$shot.follows_dribble <- ifelse(is.na(cleanshotsdf$shot.follows_dribble), 0, 1)

cleanshotsdf$shot.results.in.goal <- ifelse((cleanshotsdf$shot.outcome.name == "Goal"), 1, 0)

cleanshotsdf
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Create new variables of possible interest

cleanshotsdf$AngleToGoalNorm <- abs(cleanshotsdf$AngleToGoal - 90)

cleanshotsdf$shotnumber <- c(1:length(cleanshotsdf$shot.aerial_won))
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Remove rows with NA values

cleanshotsdf <- cleanshotsdf[(rowSums(is.na(cleanshotsdf))==0),]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Calculate variable correlations with goals scored

library(corrplot)

M = cor(cleanshotsdf[-c(1, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 16, 17, 18, 19, 20, 57)])
#corrplot(M, method = 'circle', order = "AOE", diag = FALSE, tl.pos = "lt", tl.col = "black", tl.offset=1, tl.srt = 45)
M <- as.data.frame(M)

cor_with_goal <- as.data.frame(t(as.data.frame(t(M["shot.results.in.goal"]))[-c(39)]))

friendlydf <- as.data.frame(t(as.data.frame(t(arrange(cor_with_goal, shot.results.in.goal)))[c(1:5,35:39)]))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
# Display the highest and lowest correlations with goals scored

friendlydf$row_names <- factor(row.names(friendlydf), levels = row.names(friendlydf))
ggplot(friendlydf, aes(x = row_names, y = shot.results.in.goal)) +
  geom_bar(stat = "identity", fill = "cornflowerblue") +
  labs(title = "Highest and Lowest Correlations with Goal",
       y = "Correlation with Goals") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
               plot.title = element_text(hjust=0.5),
              axis.title.x = element_blank())
```


## Data Source
### Description of match data from StatsBomb
### Overview of features within the data
### Note the unbalanced nature of the data and what we will look for to ensure it does not negatively impact our results (High AUC, sum of xG roughly equals total goals scored)

## Feature Engineering
### Shot angle to norm (we want to know how much the angle deviates from directly in front of goal)
### One hot encoding (preparing categorical variables for modeling)

## Exploring Variables of Interest
### Examining correlation with goals


# Model Fitting

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Split the dataset into training and test sets

set.seed(12345)
library(caTools)

xyz = sample.split(cleanshotsdf$under_pressure, SplitRatio = (4/5))
trainshotsdf = cleanshotsdf[xyz,]
testshotsdf = cleanshotsdf[!xyz,]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Build a full logistic regression model

mod1 = glm(shot.results.in.goal ~ under_pressure + shot.first_time + shot.one_on_one + shot.aerial_won + 
             shot.open_goal + shot.follows_dribble + DistToGoal + DistToKeeper + AngleToGoal + AngleToKeeper + AngleDeviation +
             DistSGK + density + density.incone + distance.ToD1 + distance.ToD2 + AttackersBehindBall + DefendersBehindBall +
             DefendersInCone + InCone.GK + DefArea + distance.ToD1.360 + distance.ToD2.360 + 
             shot.technique.halfvolley + shot.technique.backheel + shot.technique.volley + 
             shot.technique.overheadkick + shot.technique.divingheader + shot.technique.lob + 
             shot.body_part.head + shot.body_part.other + AngleToGoalNorm, 
           data = trainshotsdf, 
           family ="binomial")
print(summary(mod1))
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Calculate and store predicted xG from full logistic regression model
fulllogpredictedxg <- predict(mod1, newdata = testshotsdf, type = "response")

predictedxgdf <- cbind(testshotsdf$shotnumber, testshotsdf$shot.statsbomb_xg)
colnames(predictedxgdf) <- c("Shot Number", "StatsBomb xG")
predictedxgdf <- cbind(predictedxgdf, fulllogpredictedxg)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE, fig.show='hide'}
# LASSO Regression
set.seed(12345)
library(glmnet)
cv.out = cv.glmnet(as.matrix((trainshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)])), as.matrix(trainshotsdf[55]), alpha = 1, 
                   folds = 5, family="binomial")
print(cv.out$cvm)
plot(cv.out)
cat("CV Errors", cv.out$cvm,fill=TRUE)
cat("Lambda with smallest CV Error",
cv.out$lambda[which.min(cv.out$cvm)],fill=TRUE)
cat("Coefficients", as.numeric(coef(cv.out)),fill=TRUE)
cat("Number of Zero Coefficients",sum(abs(coef(cv.out))<1e-8),
fill=TRUE)

coef(cv.out)


best_lambda = cv.out$lambda[which.min(cv.out$cvm)]
best_lambda
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
set.seed(12345)
lassopredictedxg <- predict(cv.out, s = best_lambda, newx = as.matrix(testshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)]), 
                            type = "response")

predictedxgdf <- cbind(predictedxgdf, lassopredictedxg)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE, fig.show='hide'}
# Ridge Regression
set.seed(12345)
library(glmnet)
cv.out = cv.glmnet(as.matrix((trainshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)])), as.matrix(trainshotsdf[55]), alpha = 0, 
                   folds = 5, family="binomial")
print(cv.out$cvm)
plot(cv.out)
cat("CV Errors", cv.out$cvm,fill=TRUE)
cat("Lambda with smallest CV Error",
cv.out$lambda[which.min(cv.out$cvm)],fill=TRUE)
cat("Coefficients", as.numeric(coef(cv.out)),fill=TRUE)
cat("Number of Zero Coefficients",sum(abs(coef(cv.out))<1e-8),
fill=TRUE)

coef(cv.out)


best_lambda = cv.out$lambda[which.min(cv.out$cvm)]
best_lambda
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
set.seed(12345)
ridgepredictedxg <- predict(cv.out, s = best_lambda, newx = as.matrix(testshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)]), 
                            type = "response")

predictedxgdf <- cbind(predictedxgdf, ridgepredictedxg)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE, fig.show='hide'}
# Elastic Net Regression
set.seed(12345)
library(glmnet)
cv.out = cv.glmnet(as.matrix((trainshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)])), as.matrix(trainshotsdf[55]), alpha = 0.5, 
                   folds = 5, family="binomial")
print(cv.out$cvm)
plot(cv.out)
cat("CV Errors", cv.out$cvm,fill=TRUE)
cat("Lambda with smallest CV Error",
cv.out$lambda[which.min(cv.out$cvm)],fill=TRUE)
cat("Coefficients", as.numeric(coef(cv.out)),fill=TRUE)
cat("Number of Zero Coefficients",sum(abs(coef(cv.out))<1e-8),
fill=TRUE)

coef(cv.out)


best_lambda = cv.out$lambda[which.min(cv.out$cvm)]
best_lambda
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
set.seed(12345)
elasticnetpredictedxg <- predict(cv.out, s = best_lambda, newx = as.matrix(testshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)]), 
                            type = "response")

predictedxgdf <- cbind(predictedxgdf, elasticnetpredictedxg)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Random Forest

library(ISLR)
library(ISLR2)
library(tree)
library(randomForest)
library(gbm)

set.seed(12345)
p <- 32

xg.bag <- randomForest(factor(shot.results.in.goal) ~ under_pressure + shot.first_time + shot.one_on_one + shot.aerial_won + 
             shot.open_goal + shot.follows_dribble + DistToGoal + DistToKeeper + AngleToGoal + AngleToKeeper + AngleDeviation +
             DistSGK + density + density.incone + distance.ToD1 + distance.ToD2 + AttackersBehindBall + DefendersBehindBall +
             DefendersInCone + InCone.GK + DefArea + distance.ToD1.360 + distance.ToD2.360 + 
             shot.technique.halfvolley + shot.technique.backheel + shot.technique.volley + 
             shot.technique.overheadkick + shot.technique.divingheader + shot.technique.lob + 
             shot.body_part.head + shot.body_part.other + AngleToGoalNorm, 
             data = trainshotsdf,
             mtry = p,
             importance = T)

baggingpredictedxg <- predict(xg.bag, newdata = testshotsdf, type = "prob")
baggingpredictedxg <- baggingpredictedxg[,2]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
predictedxgdf <- cbind(predictedxgdf, baggingpredictedxg)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
set.seed(12345)
xg.rf1 <- randomForest(factor(shot.results.in.goal) ~ under_pressure + shot.first_time + shot.one_on_one + shot.aerial_won + 
             shot.open_goal + shot.follows_dribble + DistToGoal + DistToKeeper + AngleToGoal + AngleToKeeper + AngleDeviation +
             DistSGK + density + density.incone + distance.ToD1 + distance.ToD2 + AttackersBehindBall + DefendersBehindBall +
             DefendersInCone + InCone.GK + DefArea + distance.ToD1.360 + distance.ToD2.360 + 
             shot.technique.halfvolley + shot.technique.backheel + shot.technique.volley + 
             shot.technique.overheadkick + shot.technique.divingheader + shot.technique.lob + 
             shot.body_part.head + shot.body_part.other + AngleToGoalNorm, 
             data = trainshotsdf,
             mtry = (p^0.5),
             importance = T)

rf1predictedxg <- predict(xg.rf1, newdata = testshotsdf, type = "prob")
rf1predictedxg <- rf1predictedxg[,2]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
predictedxgdf <- cbind(predictedxgdf, rf1predictedxg)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
set.seed(12345)
xg.rf2 <- randomForest(factor(shot.results.in.goal) ~ under_pressure + shot.first_time + shot.one_on_one + shot.aerial_won + 
             shot.open_goal + shot.follows_dribble + DistToGoal + DistToKeeper + AngleToGoal + AngleToKeeper + AngleDeviation +
             DistSGK + density + density.incone + distance.ToD1 + distance.ToD2 + AttackersBehindBall + DefendersBehindBall +
             DefendersInCone + InCone.GK + DefArea + distance.ToD1.360 + distance.ToD2.360 + 
             shot.technique.halfvolley + shot.technique.backheel + shot.technique.volley + 
             shot.technique.overheadkick + shot.technique.divingheader + shot.technique.lob + 
             shot.body_part.head + shot.body_part.other + AngleToGoalNorm, 
             data = trainshotsdf,
             mtry = (p/3),
             importance = T)

rf2predictedxg <- predict(xg.rf2, newdata = testshotsdf, type = "prob")
rf2predictedxg <- rf2predictedxg[,2]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
predictedxgdf <- cbind(predictedxgdf, rf2predictedxg)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# XGBoost

library(xgboost)
library(caret)
library(tidyverse)
library(mlr)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
set.seed(12345)
params_booster <- list(booster = 'gbtree', eta = 0.3, gamma = 0, max.depth = 4, subsample = 1, colsample_bytree = 1, min_child_weight = 1,
                       objective = "binary:logistic")


bst.cv <- xgb.cv(data = as.matrix(trainshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)]), 
              label = as.matrix(trainshotsdf[c(55)]), 
              params = params_booster,
              nrounds = 100, 
              nfold = 5,
              print_every_n = 10,
              verbose = 2)

best_nrounds <- which.min(bst.cv$evaluation_log$test_logloss_mean)
best_nrounds
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
set.seed(12345)
xgboostmod <- xgboost(data = as.matrix(trainshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)]), 
                     label = as.matrix(trainshotsdf[c(55)]), 
                     nrounds = best_nrounds, params = params_booster)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
xgboostpredictedxg <- predict(xgboostmod, as.matrix(testshotsdf[c(2,13:15,21,22,27:43,45:47,48:50,53,54,56)]))
predictedxgdf <- cbind(predictedxgdf, xgboostpredictedxg)
```

## Splitting of Data
### 80/20 training/test split

## Fitting Models
### Full Logistic Regression Model
#### We fit a logistic regression model that contains all shot related features, with no variable selection
### LASSO, Ridge, and Elastic Net Logistic Regression Models
#### We fit these penalized regressions in order to prevent model overfitting
### Random Forest Models
#### We fit these models with hope that in sacrificing model interpretability, we could achieve predictive accuracy
### XGBoost Model
#### We wanted to fit an XGBoost model to the data as it is an acclaimed model that seems to boast high predictive capability.


# Model Evaluation

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
colnames(predictedxgdf) <- c("Shot.Number", "StatsBomb.xG", "Full.Logistic.Regression.xG", "LASSO.xG", "Ridge.xG",
                             "Elastic.Net.xG", "Bagging.xG", "Random.Forest.1.xG", "Random.Forest.2.xG", "XGBoost.xG")
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
library(pROC)

evaldf <- as.data.frame(predictedxgdf)[-c(1)]
evalplusdf <- cbind(evaldf, testshotsdf$shot.results.in.goal)

statsdf <- data.frame(matrix(ncol = 6, nrow = 9))
colnames(statsdf) <- c("AUC", "Specificity", "Sensitivity", "Balanced Accuracy", "Sum Total of xG", "Total xG minus Goals")
rownames(statsdf) <- c("StatsBomb.xG", "Full.Logistic.Regression.xG", "LASSO.xG", "Ridge.xG",
                             "Elastic.Net.xG", "Bagging.xG", "Random.Forest.1.xG", "Random.Forest.2.xG", "XGBoost.xG")
statsdf <- as.data.frame(t(statsdf))


# 1
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, StatsBomb.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[1] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[1])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[1] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)

# 2
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, Full.Logistic.Regression.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[2] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[2])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[2] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)

# 3
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, LASSO.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[3] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[3])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[3] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)

# 4
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, Ridge.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[4] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[4])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[4] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)

# 5
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, Elastic.Net.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[5] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[5])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[5] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)

# 6
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, Bagging.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[6] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[6])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[6] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)

# 7
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, Random.Forest.1.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[7] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[7])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[7] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)

# 8
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, Random.Forest.2.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[8] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[8])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[8] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)

# 9
roc_obj <- roc(data = evalplusdf , testshotsdf$shot.results.in.goal, XGBoost.xG)
calc_auc <- auc(roc_obj)
wxyz <- ifelse((evaldf[9] > 0.5), 1, 0)
abcd <- testshotsdf$shot.results.in.goal
pdq <- confusionMatrix(factor(wxyz), factor(abcd), mode = "everything", positive = "1")$byClass
calc_bal_acc <- pdq["Balanced Accuracy"]
calc_spec <- pdq["Specificity"]
calc_sens <- pdq["Sensitivity"]
sum_xg <- sum(evaldf[9])
sum_xg_less_goals <- sum_xg - sum(abcd)
statsdf[9] <- c(calc_auc, calc_spec, calc_sens, calc_bal_acc, sum_xg, sum_xg_less_goals)
```




```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
library(kableExtra)

as.data.frame(t(statsdf)) %>%
  kable(digits = c(3, 3, 3, 3, 1, 1), caption = "Model Evaluation Statistics", booktabs=T) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, latex_options="scale_down")
```







## Diagnostic Statistics
### AUC
### F1-Score
### Accuracy, Type I Error Rate, and Type II Error Rate
### Sum total of xG vs Sum total goals scored

## Comparison to StatsBomb xG
### Performs just worse by AUC, F1-Score criteria


# xG Model Uses

## xG Heatmap

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
library(tidyverse)

heatmap = cleanshotsdf %>% mutate(location.x = ifelse(location.x>120, 120, location.x),
location.y = ifelse(location.y>80, 80, location.y),
location.x = ifelse(location.x<0, 0, location.x),
location.y = ifelse(location.y<0, 0, location.y))
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
heatmap$xbin <- cut(heatmap$location.x, breaks = seq(from=0, to=120, by = 2),include.lowest=TRUE) 
heatmap$ybin <- cut(heatmap$location.y, breaks = seq(from=0, to=80, by = 2),include.lowest=TRUE) 
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
heatmap = heatmap%>%
group_by(xbin, ybin) %>%
mutate(xG = mean(shot.statsbomb_xg))
```


```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
library(grid)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
shotmapxgcolors <- c("#ffffff", "#f4ef95", "#FDE960", "#FCDC5F", "#F5B94D", "#F0983E", "#ED8A37", "#E66424", "#D54F1B", "#DC2608", "#BF0000", "#7F0000", "#5F0000")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
ggplot(data= heatmap, aes(x = location.x, y = location.y, fill = xG, group =xG)) + 
    geom_bin2d(binwidth = c(2, 2), position = "identity", alpha = 0.9) +
    annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) + annotate("rect",xmin = 0, xmax =
   60,ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) + annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) + annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) + annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) + annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) + annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) + annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) + annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "black", size = 0.6)+ annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+ annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
theme(rect = element_blank(), line = element_blank()) +
annotate("point", x = 12 , y = 40, colour = "black", size = 1.05) + annotate("point", x = 108 , y = 40, colour = "black", size = 1.05) + annotate("path", colour = "black", size = 0.6,
x=60+10*cos(seq(0,2*pi,length.out=2000)),
y=40+10*sin(seq(0,2*pi,length.out=2000)))+
annotate("point", x = 60 , y = 40, colour = "black", size = 1.05) + annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6, y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") + annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6, y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
  scale_fill_gradientn(colours = shotmapxgcolors) + 
  theme(axis.text.x=element_blank(),
axis.text.y=element_blank(),
axis.title.x = element_blank(),
axis.title.y = element_blank(),
plot.title = element_text(hjust=0.5),
plot.subtitle = element_text(hjust=0.5)) +
labs(title = "xG by Area on Pitch", subtitle = "La Liga, 2014/15")
```

## Player Analysis - Leo Messi xG vs Goals Scored

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
# Pull La Liga Match Data
Comp <- FreeCompetitions()[c(42),]
Matches <- FreeMatches(Comp)
StatsBombData <- free_allevents(MatchesDF = Matches, Parallel = T)
StatsBombData = allclean(StatsBombData)
messidata = StatsBombData
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
messidatadf = messidata[messidata$type.id==16,]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
cleanmessidatadf = messidatadf[c("location", "under_pressure", "type.name", "possession_team.name", "play_pattern.name", "team.id", "team.name", "player.id", "player.name", "position.name", "shot.statsbomb_xg", "shot.freeze_frame", "shot.first_time", "shot.one_on_one", "shot.aerial_won", "shot.technique.name", "shot.body_part.name", "shot.outcome.name", "shot.type.name", "match_id", "shot.open_goal", "shot.follows_dribble", "location.x", "location.y", "location.x.GK", "location.y.GK", "DistToGoal", "DistToKeeper", "AngleToGoal", "AngleToKeeper", "AngleDeviation", "DistSGK", "density", "density.incone", "distance.ToD1", "distance.ToD2", "AttackersBehindBall", "DefendersBehindBall", "DefendersInCone", "InCone.GK", "DefArea", "distance.ToD1.360", "distance.ToD2.360")]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
cleanmessidatadf <- cleanmessidatadf[cleanmessidatadf$shot.type.name=="Open Play",]
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
cleanmessidatadf$under_pressure <- ifelse(is.na(cleanmessidatadf$under_pressure), 0, 1)

cleanmessidatadf$shot.first_time <- ifelse(is.na(cleanmessidatadf$shot.first_time), 0, 1)

cleanmessidatadf$shot.one_on_one <- ifelse(is.na(cleanmessidatadf$shot.one_on_one), 0, 1)

cleanmessidatadf$shot.aerial_won <- ifelse(is.na(cleanmessidatadf$shot.aerial_won), 0, 1)

cleanmessidatadf$shot.technique.normal <- ifelse((cleanmessidatadf$shot.technique.name == "Normal"), 1, 0)
cleanmessidatadf$shot.technique.halfvolley <- ifelse((cleanmessidatadf$shot.technique.name == "Half Volley"), 1, 0)
cleanmessidatadf$shot.technique.backheel <- ifelse((cleanmessidatadf$shot.technique.name == "Backheel"), 1, 0)
cleanmessidatadf$shot.technique.volley <- ifelse((cleanmessidatadf$shot.technique.name == "Volley"), 1, 0)
cleanmessidatadf$shot.technique.overheadkick <- ifelse((cleanmessidatadf$shot.technique.name == "Overhead Kick"), 1, 0)
cleanmessidatadf$shot.technique.divingheader <- ifelse((cleanmessidatadf$shot.technique.name == "Diving Header"), 1, 0)
cleanmessidatadf$shot.technique.lob <- ifelse((cleanmessidatadf$shot.technique.name == "Lob"), 1, 0)


cleanmessidatadf$shot.body_part.leftfoot <- ifelse((cleanmessidatadf$shot.body_part.name == "Left Foot"), 1, 0)
cleanmessidatadf$shot.body_part.rightfoot <- ifelse((cleanmessidatadf$shot.body_part.name == "Right Foot"), 1, 0)
cleanmessidatadf$shot.body_part.head <- ifelse((cleanmessidatadf$shot.body_part.name == "Head"), 1, 0)
cleanmessidatadf$shot.body_part.other <- ifelse((cleanmessidatadf$shot.body_part.name == "Other"), 1, 0)


cleanmessidatadf$shot.open_goal <- ifelse(is.na(cleanmessidatadf$shot.open_goal), 0, 1)

cleanmessidatadf$shot.follows_dribble <- ifelse(is.na(cleanmessidatadf$shot.follows_dribble), 0, 1)

cleanmessidatadf$shot.results.in.goal <- ifelse((cleanmessidatadf$shot.outcome.name == "Goal"), 1, 0)
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
cleanmessidatadf$AngleToGoalNorm <- abs(cleanmessidatadf$AngleToGoal - 90)

cleanmessidatadf$shotnumber <- c(1:length(cleanmessidatadf$shot.aerial_won))
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
playerdf <- as.data.frame(unique(cleanmessidatadf["player.name"]))
playerdf$goals = 0
playerdf$xG = 0

for (i in 1:length(playerdf$player.name)) {
  playerdf$goals[i] = sum(cleanmessidatadf[(cleanmessidatadf["player.name"] == playerdf$player.name[i]),]["shot.results.in.goal"])
  playerdf$xG[i] = sum(cleanmessidatadf[(cleanmessidatadf["player.name"] == playerdf$player.name[i]),]["shot.statsbomb_xg"])
}
```

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
library(gridExtra)
library(scales)
library(colorspace)


stats <- data.frame(Category = c("Goals", "xG"), Value = c(36, 22.8))
bar_colors <- c("#EDBB00", lighten("#EDBB00", amount = 0.5))
p1 <- ggplot(stats, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity", width = 0.5, fill = bar_colors) +
  labs(title = "Lionel Messi",
       y = "Count",
       x = element_blank()) +
  theme_minimal() +
  theme(legend.position = "none",
  plot.title = element_text(hjust=0.5)) +
  ylim(c(0,40))

stats <- data.frame(Category = c("Goals", "xG"), Value = c(20, 15.8))
bar_colors <- c("#004D98", lighten("#004D98", amount = 0.5))
p2 <- ggplot(stats, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity", width = 0.5, fill = bar_colors) +
  labs(title = "Neymar",
       y = "Count",
       x = element_blank()) +
  theme_minimal() +
  theme(legend.position = "none",
  plot.title = element_text(hjust=0.5)) +
  ylim(c(0,40))

stats <- data.frame(Category = c("Goals", "xG"), Value = c(16, 12.2))
bar_colors <- c("#A50044", lighten("#A50044", amount = 0.5))
p3 <- ggplot(stats, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity", width = 0.5, fill = bar_colors) +
  labs(title = "Luis Suarez",
       y = "Count",
       x = element_blank()) +
  theme_minimal() +
  theme(legend.position = "none",
  plot.title = element_text(hjust=0.5)) +
  ylim(c(0,40))
```


```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
stats <- data.frame(Category = c("Messi", "Neymar", "Suarez"), Value = c(36/22.8, 20/15.8, 16/12.2))
bar_colors <- c("#EDBB00", "#004D98", "#A50044")
p4 <- ggplot(stats, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity", width = 0.5, fill = bar_colors) +
  labs(title = "Goal to xG Ratio",
       y = "Ratio",
       x = element_blank()) +
  theme_minimal() +
  theme(legend.position = "none",
  plot.title = element_text(hjust=0.5))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
lay <- rbind(c(1,2,3),
             c(4,4,4))
grid.arrange(p1, p2, p3, p4, layout_matrix = lay)
```




# Recap

## Overview of Results
## Limits of study and possible areas of future exploration


